---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
if (!file.exists("./data")) {
        dir.create("./data")     
}
URL <- "https://github.com/Mitridathes/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
path1 <- "./data/activity.zip"
if (!file.exists(path1)) {
        download.file(URL, path1, mode = "wb")
        unzip(path1, exdir = "./data")
}
path2 <- "./data/activity.csv"
activity <- read.csv(path2, na.strings = "NA", stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?

### Total, mean and median of the steps taken per day:
```{r}
dailyactivity <- dplyr::summarise(dplyr::group_by(activity, date),
                             totalsteps = sum(steps),
                             meansteps = mean(steps, na.rm = TRUE),
                             mediansteps = median(steps, na.rm = TRUE))
print.data.frame(dailyactivity)
```
### Histogram of the total number of steps taken each day:

```{r}
day <- format(dailyactivity$date, "%d")
with(dailyactivity, barplot(totalsteps, names.arg = day, xlab = "Day", 
                            ylab = "Total Steps", space = 0))
```

## What is the average daily activity pattern?

```{r}
intervalmean <- dplyr::summarise(dplyr::group_by(activity, interval),
                                 intmean = mean(steps, na.rm = TRUE))
with(intervalmean, plot(interval, intmean, type = "l", xlab = "Interval",
                        ylab = "Interval mean of steps across all days"))
```

### The 5-minute interval on average of all days which contains the maxium number of steps:

```{r}
intervalmean[which.max(intervalmean$intmean), ]
```

## Imputing missing values

### Calculating the number of NA's in step variable:

```{r}
sum(is.na(activity$steps))
```

### Filling in all missing values using 5 minute-interval means in a new dataset:

```{r}
fill <- function(row) {
        intervalmean[intervalmean[, 1] == row[1, 3], 2]
}
nafreeactivity <- activity
for (i in 1:nrow(activity)) {
        if (is.na(activity[i, 1])) {
                nafreeactivity[i, 1] <- fill(activity[i, ])
        }
}
```

### Histogram of the total number of steps taken each day with no NA data:

```{r}
nonadailyactivity <- dplyr::summarise(dplyr::group_by(nafreeactivity, date),
                                      totalsteps = sum(steps),
                                      meansteps = mean(steps, na.rm = TRUE),
                                      mediansteps = median(steps, na.rm = TRUE))
```

### Total, mean and median of the steps taken per day with no NA data:

```{r}
print.data.frame(nonadailyactivity)
```

### Impact of imputting data on the stimates of the total daily number of steps:

```{r}
summary(activity$steps)
summary(nafreeactivity$steps)
```

Filling NA data with 5 minute-interval means modifies the 3rd Quantile value.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# Creating a new factor for weekdays and weekends:
weekends1 <- c("sábado", "domingo")
activity$weekd <- factor((weekdays(activity$date) %in% weekends1),
                              levels = c(FALSE, TRUE),
                              labels = c("weekday", "weekend"))
# Calculating the mean of steps by interval and weekend or weekday:
intervalmean2 <- dplyr::summarise(dplyr::group_by(activity, interval, weekd),
                                  intmean2 = mean(steps, na.rm = TRUE))
# Plotting data:
library(lattice)
xyplot(intmean2 ~ interval | weekd, data = intervalmean2, type = "l",
       layout = c(1, 2), xlab = "Interval", ylab = "Average of steps taken")
```